---
title: "Dargebotene Hand"
subtitle: "Data Exploration"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

# knitr options

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# load libraries

library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(readr)
#library(statR)
library(stringr)

# ggplot theme and options

#theme_set(theme_stat(base_size = 11))


```

# Raw data input

## Shared raw data

Incoming raw data is stored in ../data/raw/.. and a new directory is created for each time that data is shared. The name of the directory is the date written in ISO 8601.

### ../data/raw/combined/..

Create a directory that will contain all shared raw data. 

```{r create-dir}

## define subdir 

sub_dir <- "data/raw_data/combined"

## define output dir

output_dir <- file.path(here::here(""), sub_dir)

## create dir if it doesnt exist

if (!dir.exists(output_dir)){
  dir.create(output_dir)
} else {
  print("Directory already exists!")
}
```

## Organise raw data

### 2020-05-28

```{r 2020-05-28-organise}
# Unzip Data 2020-05-28

unzip(zipfile = here::here("data/raw_data/2020-05-28/Kantonales_Amt_für_Statistik.zip"), exdir = here::here(sub_dir), junkpaths = TRUE)

## rename file where file name is not conform with remaining files

file.rename(
  from = file.path(output_dir, "Statistik_Stelle - 2019_2Q_nicht_bestimmbar.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2019-2Q_nicht_bestimmbar.xlsx")
) 

file.rename(
  from = file.path(output_dir, "Statistik_Stelle - 2019-1Q_Erster_Kontakt.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2019-1Q_Erstkontakt.xlsx")
)

file.rename(
  from = file.path(output_dir, "Statistik_Stelle - 2019-1Q_gelegentlicher_erster_Kontakt.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2019-1Q_gelegentlicher_Kontakt.xlsx")
)

file.rename(
  from = file.path(output_dir, "Statistik_Stelle - 2020-1Q_gelegenticher_Kontakt.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2020-1Q_gelegentlicher_Kontakt.xlsx")
)

# delete files that were incorrect see issue #2

files_del_vec <- list.files(path = here::here(sub_dir), pattern = "2019-1Q")

files_del_vec <- files_del_vec[
  !str_detect(files_del_vec, pattern = "Erst|gelegent|regel")
]


for (name in files_del_vec) {
  if (file.exists(paste0(here::here(sub_dir), "/", name))) {
    unlink(
      paste0(here::here(sub_dir), "/", name)
    )
  } else {
    
  }
  
} 

# delete files that we arent going to use

files_del_vec <- list.files(path = here::here(sub_dir), pattern = "2020-April")

for (name in files_del_vec) {
  if (file.exists(paste0(here::here(sub_dir), "/", name))) {
    unlink(
      paste0(here::here(sub_dir), "/", name)
    )
  } else {
    
  }
  
}
```

### 2020-07-13

```{r 2020-07-13-organise}

## copy files that were newly sent on 2020-07-13 into main folder

files_copy_vec <- list.files(path = here::here("data/raw_data/2020-07-13/"), pattern = "Statistik")

for (name in files_copy_vec) {
  if (file.exists(paste0(here::here("data/raw_data/2020-07-13/"), "/", name))) {
    
    file.copy(
      from = paste0(here::here("data/raw_data/2020-07-13/"), "/", name),
      to = (paste0(here::here(sub_dir), "/", name))
    )
  } else {
    
  }
}


```

### 2020-12-04

```{r 2020-12-04-organise}

raw_date_20201204 <- "2020-12-04"
path_20201204 <- str_c(here::here("data/raw_data"),"/", raw_date_20201204)

## copy files that were newly sent on 2020-07-13 into main folder

files_copy_vec <- list.files(path = path_20201204, pattern = "Statistik")

for (name in files_copy_vec) {
  if (file.exists(paste0(path_20201204, "/", name))) {
    
    file.copy(
      from = paste0(path_20201204, "/", name),
      to = (paste0(here::here(sub_dir), "/", name))
    )
  } else {
    
  }
}

## rename file where file name is not conform with remaining files

file.rename(
  from = file.path(output_dir, "Statistik_Stelle - Juli 2020.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2020-07.xlsx")
) 


file.rename(
  from = file.path(output_dir, "Statistik_Stelle - August 2020.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2020-08.xlsx")
) 


file.rename(
  from = file.path(output_dir, "Statistik_Stelle - September 2020.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2020-09.xlsx")
) 


file.rename(
  from = file.path(output_dir, "Statistik_Stelle - Oktober 2020.xlsx"),
  to = file.path(output_dir, "Statistik_Stelle - 2020-10.xlsx")
) 



```

## Import raw data

```{r}

no_files <- length(list.files(path = here::here(sub_dir)))

dat_in <- list()

for (i in seq(1:no_files)) {
  dat_in[[i]] <- readxl::read_excel(
    path = here::here(sub_dir, list.files(path = here::here(sub_dir))[[i]]),
    
    ## Skip the first two rows
    
    skip = 2
  ) 
}

```

# Data tidying

Preapere clean variables and rename unnamed variables. All rows with only NAs are removed from the data frames and the variable "gruppe" filled to have a complete tidy dataframe. 

```{r}

dat_clean <- list()

## clean dataframe

for (i in seq(1:no_files)) {
  
  dat_clean[[i]] <- dat_in[[i]] %>% 
    janitor::clean_names() %>% # View()
    rename(
      gruppe = x1,
      variable = x2
    ) %>% 
    filter_all(any_vars(!is.na(.))) %>%  ## TODO: add to glossary filter out all rows with only NAs 
    slice(-1) %>% 
    fill(gruppe, .direction = "down") 
  
}

```

## Validation 1: Check sums (good)

The third row of the raw data files contains the total number of calls in column `Anzahl`. This number is compared to the sum of calls in variable `Gruppe` for value 'Schicht', as this is the only value that is recorded for each incoming call (e.g. Alter, Geschlecht are not always recorded).

```{r check-1}
# Generate random file number
random_file_no <- sample(seq(1:no_files), 1)

# Extract value from raw data file for a randon file
anzahl_anrufe_total <- dat_in[[random_file_no]] %>% slice(1) %>% .$Anzahl

# Write check
check1 <- dat_clean[[random_file_no]] %>%
  filter(gruppe == "Schicht") %>% 
  group_by(gruppe) %>% 
  summarise(
    sum_anzahl = sum(anzahl)
  ) %>% 
  mutate(
    check = anzahl_anrufe_total == sum_anzahl
  )

# Stop script if check is FALSE
if (check1$check == FALSE) {
  print("Check 1 of Validation 1 is FALSE")
  stop("Check 1 of Validation 1 is FALSE. Please evaluate error in code chunk 'check-1' of 'clean_raw_data.Rmd")
  
} else {
  print("Check 1 of Validation 1 is TRUE")
}


```

## Add metadata 

Relevant metadata is stored in the file names. This metadata is extracted.

```{r metadata-extract}
## use filenames to identify grouped variable year and qaurtal
dat_filename <- tibble(
  file_id = seq(1:no_files),
  file_name = list.files(path = here::here(sub_dir))
)

## filename dataframe
dat_filename_df <- dat_filename %>% 
  separate(col = file_name, into = c("a", "b", "c", "d"), sep = c(19, 24, 26)) %>% 
  separate(col = d, into = c("e", "f"), sep = ".xlsx") %>% 
  select(-a, -f) %>% 
  mutate(jahr = str_replace(b, pattern = "-", replacement = "")) %>% 
  select(-b) %>% 
  mutate(
    quartal = case_when(
      str_detect(c, pattern = "Q") ~ c
    )
  ) %>% 
  mutate(
    monat = case_when(
      c == "Ap" ~ "04",
      is.na(quartal) ~ c
    )
  ) %>% 
  select(-c) %>% 
  rename(aggregiert = e) %>% 
  mutate(jahr = str_replace(jahr, "_", ""))

## replace "" with NA
dat_filename_df$aggregiert <- na_if(dat_filename_df$aggregiert, "")

```

### Aggregation level

The metadata of the file names allows to add variables for the level of aggregation of each dataframe.

```{r}

# join gruppe, jahr, quartal and monat to individual tables

dat_final <- list()
sliced <- list()

for (i in seq(1:no_files)) {
  
  sliced[[i]] <- slice(dat_filename_df, i)
  
  dat_final[[i]] <- dat_clean[[i]] %>% 
    mutate(
      file_id = sliced[[i]]$file_id,
      aggregiert = sliced[[i]]$aggregiert,
      jahr = sliced[[i]]$jahr,
      quartal = sliced[[i]]$quartal,
      monat = as.numeric(sliced[[i]]$monat)
    ) %>% 
    mutate(
      aggregiert = case_when(
        aggregiert == "-18" ~ "bis 18",
        aggregiert == "_19-40" ~ "19 bis 40",
        aggregiert == "_41-65" ~ "41 bis 65",
        aggregiert == "_65-" ~ "über 65",
        aggregiert == "_nicht_bestimmbar" ~ "nicht bestimmbar",
        aggregiert == "_nicht bestimmbar" ~ "nicht bestimmbar",
        aggregiert == "_Erstkontakt" ~ "Erstkontakt",
        aggregiert == "_gelegentlicher_Kontakt" ~ "gelegentlicher Kontakt",
        aggregiert == "_regelmässiger_Kontakt" ~ "regelmaessiger Kontakt",
        aggregiert == "_regelm„ssiger_Kontakt" ~ "regelmaessiger Kontakt",
        aggregiert == "_regem„ssiger_Kontakt" ~ "regelmaessiger Kontakt",
        TRUE ~ aggregiert
      )
    )
}

```

## Improve data strcture

As we wrote the article and wanted to publish the data openly on GitHub, we realised that some variables are not relevant for the anaysis und that the data structure could be improved. The final data cleaning step is included in the code chunk below.

### Monthly data

```{r}
dat_by_month <- list()

for (i in seq(1:no_files)) {
  dat_by_month[[i]] <- dat_final[[i]] %>% 
    filter(is.na(quartal))
}

## Final data cleaning

dat_by_month_export <- dat_by_month %>% 
  bind_rows() %>% 
  filter(!gruppe %in% c("Opferhilfe", "Wohnregion", "Sprache")) %>% 
  rename(kategorie = gruppe, sub_kategorie = variable, wert = anzahl) %>% 
  mutate(
    einheit = case_when(
      kategorie == "Gesprächsdauer" ~ "minuten",
      kategorie == "Beratungsinhalt" ~ "anzahl_inhalte",
      TRUE ~ "anzahl_anrufe"
    )
  ) %>% 
  select(-in_prozent, -file_id, -aggregiert, -quartal) %>% 
  relocate(einheit, .after = "wert")


```

### Quarterly data

```{r}
dat_by_quartal <- list()

for (i in seq(1:no_files)) {
  dat_by_quartal[[i]] <- dat_final[[i]] %>% 
    filter(!is.na(quartal))
}

aggregation_alter <- c("bis 18", "19 bis 40", "41 bis 65", "19 bis 65", "über 65", "nicht bestimmbar", "Total")
aggregation_kontakthaeufigkeit <- c("Erstkontakt", "gelegentlicher Kontakt", "regelmaessiger Kontakt")

dat_by_quartal_export <- dat_by_quartal %>% 
  bind_rows() %>% 
  filter(!gruppe %in% c("Opferhilfe", "Wohnregion", "Sprache")) %>% 
  rename(kategorie = gruppe, sub_kategorie = variable, wert = anzahl) %>% 
  select(-in_prozent, -file_id, -monat) %>%
  filter(aggregiert %in% aggregation_alter) %>% 
  filter(kategorie != "Alter") %>% 
  mutate(
    einheit = case_when(
      kategorie == "Gesprächsdauer" ~ "minuten",
      kategorie == "Beratungsinhalt" ~ "anzahl_inhalte",
      TRUE ~ "anzahl_anrufe"
    )
  ) %>% 
  relocate(aggregiert, .after = "sub_kategorie") %>% 
  relocate(einheit, .after = "wert") %>% 
  #relocate(sub_kategorie, .after = "kategorie")
  rename(altersklasse = aggregiert)

```

# Data Export

Write individual CSV for monthly and quarterly data.

```{r data-export}

write_csv(dat_by_month_export, here::here("data/intermediate_data/Social_Tel143_nach_monat.csv"))
write_csv(dat_by_quartal_export, here::here("data/intermediate_data/Social_Tel143_nach_quartal.csv"))

```

## Validation 2: 

```{r}

# Compare data for Q2 2019 with April, May, June 2019

dat_by_quartal_export %>% 
  filter(quartal == "2Q", jahr == "2020") %>% 
  
  ## Filter nach aggregierung
  filter(kategorie == "Schicht") %>% 
  #filter(aggregiert %in% aggregiert_alter) %>% 
  
  group_by(kategorie, altersklasse) %>%
  
  ## First summarise auf Ebene Gruppe Schicht und aggregiert
  
  summarise(
    sum_anzahl = sum(wert)
  ) 


# ## Second summarise to get total

## by month

dat_by_month_export %>% 
  filter(monat %in% c(4, 5, 6), jahr == "2020") %>% 
  #filter(gruppe == "Schicht") %>% 
  group_by(kategorie, sub_kategorie) %>% 
  summarise(
    sum_anzahl = sum(wert)
  ) %>% 
  filter(kategorie == "Alter")  
  
```

# README Markdown

```{r}

dat_by_month_export %>% 
  names() %>% 
  tibble::enframe() %>% 
  select(Spaltenname = value) %>% 
  mutate(
    Beschreibung = c(
      "Kategorien zu welchen bei einem Anruf Informationen aufgenommen werden",
      "Jeweilige Subkategorie für die einzelnen Kategorien",
      "Wert für die jeweiligen Subkategorien (siehe 'Wichtige Anmerkung' für Details zu Kategorien)",
      "Einheit des Werts",
      "Kalenderjahr",
      "Monat des Kalenderjahres")
    ) %>% 
  mutate(
    Format = c(
      "Text", "Text", "Zahl", "Text", "Zahl", "Zahl"
    )
  ) %>% 
  knitr::kable()

```

